---
title: Pitch Histograms
---

```{r, echo=F, warnings=FALSE, message=FALSE}
library(tidyverse)
library(humdrumR)
```


### Pitch histograms in Freygish

```{r, freygish-histograms, echo=F, warnings=FALSE, message=FALSE}
library(tidyverse)
freygish <- readHumdrum('corpus/extracts/freygish/*.krn')
freygish_in_G <- freygish |> mutate(Transposed = transpose(Token, from = Key, to = 'G'))
freygish_1 <- freygish_in_G |> 
    pitch(simple = FALSE) |>
    humdrumR::tally() |> as_tibble() |> 
    filter(!grepl("[CDEF].*3", Pitch)) |> filter(!grepl("[789]", Pitch)) |> 
    filter(!grepl("[A|B].*6", Pitch)) |>
    filter(!grepl("G#6", Pitch))

freygish_1$Pitch <- factor(freygish_1$Pitch, levels = freygish_1$Pitch)
    
ggplot(freygish_1, aes(x=Pitch, y=n)) +
    geom_bar(stat="identity") + theme_bw() +
    labs(y="count") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

As a percentage.

```{r, freygish-histograms-percentage, echo=F, warnings=FALSE, message=FALSE}
x <-  freygish_in_G |> 
    pitch(simple = FALSE) |>
    humdrumR::tally() 
     
percentage <- x/sum(x) * 100
percentage <- as_tibble(percentage)
percentage$Pitch <- factor(percentage$Pitch, levels = percentage$Pitch)   
percentage <- percentage |> 
    filter(!grepl("[CDEF].*3", Pitch)) |> filter(!grepl("[789]", Pitch)) |> 
    filter(!grepl("[A|B].*6", Pitch)) |>
    filter(!grepl("G#6", Pitch))

ggplot(percentage, aes(x=Pitch, y=n)) +
    geom_bar(stat="identity") + theme_bw() +
    labs(y="percentage") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

### Pitch histograms for the Raised Fourth mode.

```{r, raised-fourth-histogram, echo=F, warnings=FALSE, message=FALSE}
raised_four <- readHumdrum('corpus/extracts/raised_fourth/*.krn')
raised_four_1 <- raised_four |> 
    pitch(simple = FALSE) |>
    humdrumR::tally() |> as_tibble() |>
    filter(!grepl("[CDEF].*3", Pitch)) |> filter(!grepl("[789]", Pitch)) |> 
    filter(!grepl("[A|B].*6", Pitch)) |>
    filter(!grepl("G#6", Pitch))

raised_four_1$Pitch <- factor(raised_four_1$Pitch, levels = raised_four_1$Pitch)
    
ggplot(raised_four_1, aes(x=Pitch, y=n)) +
    geom_bar(stat="identity") + theme_bw() +
    labs(y="count") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

And again as a percentage:

```{r, raised-fourth-percentage, echo=F, warnings=FALSE, message=FALSE}
x <-  raised_four |> 
    pitch(simple = FALSE) |>
    humdrumR::tally() 
percentage <- x/sum(x) * 100
percentage <- as_tibble(percentage)
percentage <- percentage |> 
    filter(!grepl("[CDEF].*3", Pitch)) |> filter(!grepl("[789]", Pitch)) |> 
    filter(!grepl("[A|B].*6", Pitch)) |>
    filter(!grepl("G#6", Pitch))

percentage$Pitch <- factor(percentage$Pitch, levels = percentage$Pitch) 

ggplot(percentage, aes(x=Pitch, y=n)) +
    geom_bar(stat="identity") + theme_bw() +
    labs(y="percentage") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


### Pitch histograms for the Major mode.

```{r, major-histogram, echo=F, warnings=FALSE, message=FALSE}
major <- readHumdrum('corpus/extracts/major/*.krn')
major_1 <- major |> 
    pitch(simple = FALSE) |>
    humdrumR::tally() |> as_tibble() |>
    filter(!grepl("[CDEF].*3", Pitch)) |> filter(!grepl("[789]", Pitch)) |> 
    filter(!grepl("[A|B].*6", Pitch)) |>
    filter(!grepl("G#6", Pitch))


major_1$Pitch <- factor(major_1$Pitch, levels = major_1$Pitch)
    
ggplot(major_1, aes(x=Pitch, y=n)) +
    geom_bar(stat="identity") + theme_bw() +
    labs(y="count") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


Looking at this as a percentage of total pitches might be more useful. The numbers on the y-axis represent the percentage of pitches used in a given mode. 

```{r, major-percentage, echo=F, warnings=FALSE, message=FALSE}
x <-  major |> 
    pitch(simple = FALSE) |>
    humdrumR::tally() 
percentage <- x/sum(x) * 100
percentage <- as_tibble(percentage)
percentage <- percentage |> 
    filter(!grepl("[CDEF].*3", Pitch)) |> filter(!grepl("[789]", Pitch)) |> 
    filter(!grepl("[A|B].*6", Pitch)) |>
    filter(!grepl("G#6", Pitch))

percentage$Pitch <- factor(percentage$Pitch, levels = percentage$Pitch)
ggplot(percentage, aes(x=Pitch, y=n)) +
    geom_bar(stat="identity") + theme_bw() +
    labs(y="percentage") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

Note: Other major keys can be found [here](other_major_keys.qmd).

### Pitch histograms for the minor mode.

```{r, minor-histogram, echo=F, warnings=FALSE, message=FALSE}
minor <- readHumdrum('corpus/extracts/minor/*.krn')
minor_1 <- minor |> 
    pitch(simple = FALSE) |>
    humdrumR::tally() |> as_tibble() |>
    filter(!grepl("[CDEF].*3", Pitch)) |> filter(!grepl("[789]", Pitch)) |> 
    filter(!grepl("[A|B].*6", Pitch)) |>
    filter(!grepl("G#6", Pitch))

minor_1$Pitch <- factor(minor_1$Pitch, levels = minor_1$Pitch)
    
ggplot(minor_1, aes(x=Pitch, y=n)) +
    geom_bar(stat="identity") + theme_bw() +
    labs(y="count") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


```

Again, here we might explore it as a percentage of total pitches used in this mode.

```{r, minor-percentage, echo=F, warnings=FALSE, message=FALSE}
x <-  minor |> 
    pitch(simple = FALSE) |>
    humdrumR::tally() 
percentage <- x/sum(x) * 100
percentage <- as_tibble(percentage)
percentage <- percentage |> 
    filter(!grepl("[CDEF].*3", Pitch)) |> filter(!grepl("[789]", Pitch)) |> 
    filter(!grepl("[A|B].*6", Pitch)) |>
    filter(!grepl("G#6", Pitch))

percentage$Pitch <- factor(percentage$Pitch, levels = percentage$Pitch)

ggplot(percentage, aes(x=Pitch, y=n)) +
    geom_bar(stat="identity") + theme_bw() +
    labs(y="percentage") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


# The Meertens Tune Collection of Dutch Folksongs

For comparison, here we have the pitch distributions of the major and minor modes from the Meertens Tune Collection of Dutch Folksongs.

## Major

```{r, meertens-histogram-major, echo=F, warnings=FALSE, message=FALSE}
meertens_major <- readHumdrum('corpus/meertens/meertens_major/*.krn')
meertens_major_in_G <- meertens_major |> mutate(Transposed = transpose(Token, from = Key, to = 'G'))


major_1 <- meertens_major_in_G |> 
    pitch(simple = FALSE) |>
    humdrumR::tally() |> as_tibble() |>
    filter(!grepl("[GABCDEF].*3", Pitch)) |> filter(!grepl("[789]", Pitch)) |> 
    filter(!grepl("[A|B].*6", Pitch)) |>
    filter(!grepl("G#6", Pitch))


major_1$Pitch <- factor(major_1$Pitch, levels = major_1$Pitch)
    
ggplot(major_1, aes(x=Pitch, y=n)) +
    geom_bar(stat="identity") + theme_bw() +
    labs(y="count") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


## Minor


```{r, meertens-histogram-minor,echo=F, warnings=FALSE, message=FALSE}
meertens_minor <- readHumdrum('corpus/meertens/meertens_minor/*.krn')
meertens_minor_in_G <- meertens_minor |> mutate(Transposed = transpose(Token, from = Key, to = 'G'))
#meertens_minor <- rename(meertens_major, Pitch = V1, n = V2)

minor_1 <- meertens_minor_in_G |> 
    pitch(simple = FALSE) |>
    humdrumR::tally() |> as_tibble() |>
    filter(!grepl("[GABCDEF].*3", Pitch)) |> filter(!grepl("[789]", Pitch)) |> 
    filter(!grepl("[A|B].*6", Pitch)) |>
    filter(!grepl("G#6", Pitch))


minor_1$Pitch <- factor(minor_1$Pitch, levels = minor_1$Pitch)
    
ggplot(minor_1, aes(x=Pitch, y=n)) +
    geom_bar(stat="identity") + theme_bw() +
    labs(y="count") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```